The transboundary nature of the world’s exploited marine species

### 2020-10-05 Scripts/Manuscript_Draft.Rmd
Fixed indexing

---
title: "The transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes*^1^, Gabriel Reygondeau^1,2^, Colette C.C. Wabnitz^1,3,4^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Yale University, New Haven, Connecticut, USA. ^3^Stockholm Resilience Center, Stockholm University, Stockholm, Sweden. ^4^Center for Ocean Solutions, Stanford University, Stanford, United States"
# subtitle: "WORKING PAPAER. PLEASE DO NOT SHARE"
output:
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
  pdf_document:
    fig_caption: yes
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ../References/FishForVisa_Ref.bib
csl: ../References/nature_cc.csl
---

**Corresponding** author: j.palacios@oceans.ubc.ca

@@

<!-- # Instructions for submission -->

<!-- - Journal: *Scientific Reports* -->
<!-- - Information to authors (https://www.nature.com/srep/author-instructions/submission-guidelines) -->
<!-- - There are two types of article: -->

<!-- -- Articles should be no more than 11 typeset pages in length. As a guide, the main text (not including Abstract, Methods, References and figure legends) should be no more than 4,500 words. The maximum title length is 20 words. The Abstract (without heading) - which must be no more than 200 words long and contain no references - should serve both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.. -->

# Abstract

Regulatory boundaries and species distributions often do not align. This is especially the case of marine species crossing multiple Exclusive Economic Zones (EEZs). Such movements represent a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits of shared marine species. Here, we combined species distributions and the spatial delineation of EEZs to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% of the species analyzed are transboundary (n = 633). Between 2005 and 2014, fisheries targeting these species within global-EEZs caught on average 48 million tonnes per year, equivalent to an average of USD 77 billion in annual fishing revenue. For select countries, over 90% of their catch and economic benefits were attributable to a few shared resources. Our analysis suggests that catches from transboundary species are declining more than those from non-transboundary species. Our study has direct implications for managing fisheries on transboundary species, highlighting the need for strengthened effective international cooperation.

# Introduction

@@

# Results and Discussion

We identified 633 exploited transboundary species worldwide (67.5% of the 938 species analyzed), almost double previous estimates [@Teh:2015gd]. Between 2005 and 2014, national fleets targeting these transboundary species within EEZs caught an annual average of 48.5 million tonnes, representing 82.3% of EEZ-derived catches reconstructed by the *Sea Around us* at the species level (Fig. 1a). These catches generated a yearly average of USD 77,591 million in fishing revenue (78.5% of the global fishing revenue) over the same time period. Our findings are considerably higher than the previous estimates of 34.2 million tonnes in catches and USD 40,187 million (in 2019 value) in fishing revenue from shared stocks [@Teh:2015gd].When we re-estimated the transboundary species’ catches using data consistent with those used in previous studies (FAO global reported data in year 2006), the 633 transboundary species identified here accounted for 40.4 million tonnes of annual catches (i.e., 18% higher than previously estimated). Our results suggests that the contribution of transboundary species to global catch and fishing revenue might have been previously underestimated due to an incomplete understanding of the transboundary nature of marine fished species. The 305 non-transboundary species (termed as ‘discrete’ species here, see methods - Determining transboundary species trait) accounted for a much smaller proportion of total catch and revenue; 2.8 million tonnes and USD 4,282 million annually respectively, on average, between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

```{r Call_Map, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species and their contribution to global fisheries catch and revenue. a) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. b) Contribution of transboundary species to regional revenue (left) and catch (right). Regions classified according to the United Nations sub-regions. Points = mean +- sd. Revenue in 2010 real USD"}

![Figure 1. Number of transboundary species and their contribution to global fisheries catch and revenue. a) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. b) Contribution of transboundary species to regional revenue (left) and catch (right). Regions classified according to the United Nations sub-regions: E. = Eastern, N. = Northern, S. = Southern, W. = Western, Ltn. Ame. & the Car. = Latin America and the Caribbean. Points = regional  mean ± sd. Revenue in 2019 real USD. Figure created using R version 3.5.2. Land and EEZ shapefiles from Natural Earth (www.naturalearthdata.com) and Sea Around Us (www.seaaroundus.org), respectively.](../Figures/Fig1.png)

At a regional level, we found that transboundary species are particularly economically important for Northern America (average country revenue = USD 4,680 $\pm$ 6,000 million) and Eastern Asia (USD 3,779 $\pm$ 3,093 million) (Fig. 1b). The estimated per country revenues from transboundary species in these two regions is significantly different from other regions (one-way ANOVA; *DF*(16,165) = 5.081, *p* < 0.001, $\alpha$ = 0.05). China (USD 7,284 million) and the USA (USD 11,604 million) contribute 55% and 82% to the annual average revenue from 2005 to 2014  in Eastern Asia and Northern America, respectively. In addition to China and the USA, Russia (USD 7,379 million), Peru (USD 6,044 million) and Japan (USD 3,907 million) were among the top five countries with the most fishing revenue generated from transboundary species between 2005 and 2014 (Fig. 1a). These five nations were responsible for 41% of the yearly global fisheries revenue from transboundary species.

Peru and Russia, having the two largest fisheries by total catch in the world [@FAO:2018KK], each responsible for over 5.8 million tonnes annually, on average, between 2005 and 2014  (Supplementary Fig. S1). Peru's catches consist mainly of Peruvian anchoveta (*Engraulis ringens*) and account for 79% of the national transboundary species production. Peru and Chile recently signed an agreement to work towards standardized stock assessments through coordinated management of the southern anchoveta stock [@UNDP:2016wp]. A similar management agreement was signed by Russia, Japan and the USA over shared Alaskan pollock (*Theragra chalcogramma*) in the Bering Sea in 1988 [@NOAA:gX6CtUTk]. Transboundary species also make large contributions to fisheries in Eastern Asia (one-way ANOVA; *Df*(16,158) = 2.265, *p* = 0.005, $\alpha$ = 0.05). China, the world's top fish producer [@FAO:2018KK], obtains one third (5.1 million tonnes) of its total catches from transboundary species, followed by Japan (1.8 million tonnes) and South Korea (1.06 million tonnes). Differences in the regional importance of transboundary fisheries can also be illustrated in terms of catch-revenue over the area ($km^2$) of the EEZ (Fig. 2). As an example, the aggregated EEZ area for all Northern European countries where transboundary species are present was 3.3 million $km^2$, the $6^{th}$ smallest of the 17 groups analyzed (Supplementary Table S1). However, it had the second highest average revenue (USD 26.1 thousand per $km^2$) and the highest average catch (19.9 tonnes per $km^2$) of transboundary species per EEZ area between 2005 and 2014. At the country level, countries from Western Europe accrued significantly more revenue from transboundary fisheries per $km^2$ than any other country (one-way ANOVA, *Df*(16,165) = 3.267, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.05; Fig. 2).

```{r Call_Fig_2, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap=""}

![Figure 2. Weighted benefits of transboundary species by $km^{2}$ and UN sub-region. a) Revenue in thousand 2019 USD. b) Catch in tonnes. Points are sub-region averages $\pm$ s.d. by country. Figure created using R version 3.5.2. Abbreviations as in Figure 1.](../Figures/Fig2.png)

We determined the catch trend of each species within each EEZ, classifying them as increasing (Category A), constant (Category B) or decreasing (Category C) (Fig. 3). While previous studies have demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Our intention here is to examine where the catch trends of transboundary species differ from non-transboundary species between 2005 and 2010 relative to historic catch since 1951 (See methods). We found significant differences for all catch trend categories for transboundary species (one-way ANOVA, *Df*(2,459) = 47.94, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.001), and no significant differences in catch trend categories for discrete species (one-way ANOVA, *Df*(2,106) = 1.885, *p* = 0.157; $\alpha$ = 0.05). We also found significant differences in catch trends when directly comparing transboundary to discrete species categories (MANOVA, *Df*(2,459) = 19.001, *p* < 0.001). Overall, transboundary species only targeted by one country are generally less likely to have a decreasing catch trend compared to instances where the shared species is fished by neighboring countries (Supplementary Table S2).

```{r Call_Circ_Country, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species by catch category. Category A, increasing catch; Category B, constant catch; Category C, decreasing catch"}

![Figure 3. Number of transboundary species by catch trend and EEZ. Category A, Increasing; Category B, Constant; Category C, Decreasing. “No Category” reflects species with less than 10 years of catch data and/ or less than 5 consecutive years of catch data. Figure created using R version 3.5.2.](../Figures/Fig3.png)

<!-- Europe has the higest rate of transboundary species per $km^2$ (Figure 2) which is consistent with their Common Fisheries Policy to manage the EEZ as a whole. However, despite some cases of successfull fisheries management, excessive quotas, poorly implemented fishing subsidies, and policy-related issues have failed to control fishing effort threating the sustainabillity of most european stocks [@Khalilian:2010bk].  -->

Most marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk; @Reygondeau:2019uh], from highly migratory species associated with pelagic-oceanic ecosystems like tunas (*Thunnus sp.*), to less mobile reef-associated species like greater amberjack (*Seriola dumerili*), and species found in demersal ecosystems like gilthead seabream (*Sparus aurata*). Species associated to pelagic-oceanic ecosystems are the only group whose EEZ range (i.e., the number of EEZs where the species occur as transboundary) was significantly different from other groups (one-way ANOVA; *DF*(5,597) = 53.82, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species. The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these species have a narrower distribution or are less mobile (Fig. 4a). This result is likely an effect of the broad distribution that many of the species with preference for pelagic-oceanic ecosystems have. Many pelagic-oceanic species are highly migratory and straddle across EEZs while crossing the high seas. Thus, the number of non-neighboring EEZs sharing a highly migratory species can be over 100, as in the case of bigeye tuna (*Thunnus obesus*) (Fig. 4a). Due to their vast migration patterns and presence in areas beyond national jurisdiction, many highly migratory species are managed by Regional Fisheries Management Organizations (RFMOs).

```{r Call_Spp_Circular, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of countries shared by transboundary species. Showing only species that share > 20 countries"}

![Figure 4. Number of EEZs shared by transboundary species. a) Number of EEZs by species organized according to ecosystem preference as defined by FishBase. Showing only species that share > 20 EEZs. b) Average catch between 2005 and 2014 for top five countries capturing the top five shared species for each ecosystem preference category (color coding is as shown in legend in a). Note that there could be > 5 species due to similar sharing values. The category “Other” consists of species that have no ’ecosystem preference’ classification in FishBase. Figure created using R version 3.5.2.](../Figures/Fig4.png)

Many transboundary species are not considered highly migratory but are still shared by numerous neighboring EEZs (e.g. garfish *Belone belone*) (Fig. 4a). In addition, many fish stocks have meta-populations that are connected through larvae dispersal with ‘source’ populations potentially supporting ‘sink’ populations thousands of kilometers away [@Ramesh:2019va]. For example, while coral reef-related species were found to share fewer neighboring EEZs than other species (Fig. 4a), coral regions often share multiple species through larval connectivity [@Schill:2015fh] and adult movement [@Levin:2018in]. However, **it is important to acknowledge the uncertainty in the connectivity of marine populations at different life stages from larvae [@Kaplan:2016hs] to adults [@Archambault:2016ie].** In some areas like the Caribbean and the Western Indian Ocean, among other regions, transboundary marine protected areas have been identified as potential tools to support fisheries and conservation goals [@Levin:2018in; @Perez:2019dm]. The effective management of coral reef species is critical to many coastal communities, as they tend to highly dependent on subsistence fishing for food and nutrition security as well as livelihoods [@CisnerosMontemayor:2016gq; @Hanich:2018dh]. For instance, a number of countries with the largest catches of transboundary reef and pelagic-oceanic associated species (Fig. 4b) are also associated with some of the highest rates of fish consumption [@FAO:2018KK]. In the Philippines both pelagic and reef fishes contribute substantially to both local food and nutrition security as well as livelihoods [@Cabral:2018hc].

# Conclusions

# Materials and Methods

## Databases of species' geographic distribution

The *Sea Around Us* has reconstructed global fisheries catches from 1950 to 2014, identifying commercial marine fish and invertebrates as well as fishing regions [@Zeller2016]. We used the *Sea Around Us* reconstructed species list to determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs. The total number of species analyzed was 938, representing 67% of identified species by the *Sea Around Us* and accounting for 96.52% of the catch identified at the species’ level. To determine the current distribution of exploitable marine species, we used four data sources of species-distributions: (*i*) occurrence data, (*ii*) an Ensemble Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (Table 1). Each source represents a different method of estimating the distribution of a given species, thus providing a more robust result than an analysis focused on a single data source. Only commercial fished species with data from all four sources were included in our analysis.

Table 1. Summary of data sources to estimate species' distributions. All data has a resolution  $0.5^{\circ}$ latitude x $0.5^{\circ}$ longitude. All data are publicly available (see Acknowledgments).

@@

data.frame(Source,
           Abreviation,
           Main_Method,
           Sources) %>% 
  rename("Main Method"=Main_Method) %>% 
  # kable("latex", booktabs = T,
  #       caption = "Summary of data sources to estimate species' distributions. All data has a ressolution of 0.5 degrees latitude x 0.5 degrees longitude. All data are publically available, see references for details.") %>%
  # kable_styling(full_width = F) %>%
  # column_spec(2,width = "8cm") %>% #wraps text in column2
  # column_spec(3,width = "3cm") %>% #wraps text in column2
  write_csv("Table_1.csv")

### *Occurrence data*

@@

### Criteria 3; Spatial distribution

Finally, to have a more robust result and not categorize a species as transboundary based on its presence in a single $0.5^{\circ}$ x $0.5^{\circ}$ grid cell within an EEZ, we computed an *Area Index*. The *Area Index* represents the proportion of a given species' overarching shared distribution between neighboring EEZs accounted for by each individual EEZ. We classified a species as transboundary if both neighboring EEZs enclosed over 25% of the species joint shared distribution. While a species that has less than the selected threshold is not considered transboundary in this paper, this threshold can be lowered for a more relaxed result or increased for a more conservative estimate (Supplementary Fig. S2).

## Fisheries trends

@@

We used catch data as described above to determine the catch trend of each species within an EEZ. Although this method has previously been used to estimate stock status [@Grainger:1996tk], the categories presented here are intended to represent catch trends, and not fishing status for each species as many environmental and social-economic factors (e.g., temperature, markets, fishing policies, and fishing effort) affect catches [@Pauly:2013gs; @Branch:2008kk]. We only assessed species within each EEZ for which at least ten years of data were available between 1951 and 2014 and with at least five consecutive years of data. Three final categories were drawn for each species depending on catch volume within each EEZ (e.g., present, maximum, and minimum EEZ's historical catch) and the year (e.g., year of maximum historical catch of the species within that EEZ) (Table 2) [@Kleisner:2011wn]. Accordingly, Category A represents fisheries that are registering increases in catch (“increasing”); Category B, species that have a constant catch rate (“constant”); and Category C, species that have registered declines in catch over the last 10 years (“decreasing”). Finally, we reported the predominant category over the time period considered.

Table 2. Rules to determine the category of each analyzed species.

```{r Status_Table, eval = F, echo = F}

### Original version

# Stock_Status <- c("Rebuilding","Developing","Max Exploited","Over Exploited","Collapsed","No Status")

# Rules <- c(
#   "(Year of Catch > Year Post Max. Min. & Post Max Min Catch < (Max Catch*0.10)) & (Catch > (Max Catch*0.10) & Catch < (Max Catch*0.50))", # Reb.
#   "Year of Catch < Year of Max. Catch & Catch <= (Max Catch*0.50) or Year of Max Catch = Last Year of data)", #Dev.
#   "Catch > (Max Catch*0.50)", #Max Exp.
#   "Year of Catch > Year of Max. Catch & (Catch > (Max. Catch*0.10) & Catch < (Max. Catch*0.50)", # Over.
#   "Year or Catch > Year of Max. Catch & Catch < Max. Catch*0.10", # Colla.
#   "Non of the above rules applied"
# )

# data.frame(Stock_Status,
#            Rules) %>% 
#   kable("latex", booktabs = T,
#         caption = "Rules to determine the status of each transboundary species") %>%
#   kable_styling(full_width = F) %>%
#   column_spec(1,width = "3cm") %>% #wraps text in column2
#   column_spec(2,width = "13cm") #%>% #wraps text in column2

### Grouped version

@@

data.frame(Categories,
           Rules) %>% 
  # kable("latex", booktabs = T,
  #       caption = "Rules to determine the category of each transboundary species") %>%
  # kable_styling(full_width = F) %>%
  # column_spec(1,width = "3cm") %>% #wraps text in column2
  # column_spec(2,width = "13cm") %>% #wraps text in column2
  write_csv("Table_2.csv")

@@

We have identified four key uncertainties in the method utilized that may affect the estimation of transboundary species. Firstly, as we ran the analysis at the species level due to limited spatial-specific data on species sub populations (stocks), we are not able to identify transboundary stocks within EEZs. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level. Also, recent research suggests strong connectivity between stocks, even when separated by thousands of kilometers [@Ramesh:2019va] providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca]. **However, it is important to acknowledge that there is considerable uncertainty associated in determining levels of connectivity across time and space for marine populations at different life stages, from larvae [@Kaplan:2016hs] to adults [@Archambault:2016ie]. This is of special concern, but-not-limited-to [@Archambault:2016ie], highly migratory species whose ranges span multiple jurisdictions and the high seas, such as tunas [@Moore:2020cd], billfishes [@Sepulveda:2019hq] and sharks [@Vandeperre:2014jk] challenging management decisions based on meta-populations [@Moore:2020cd].** Here, we limited the definition of “transboundary” to include species spanning only adjacent countries (e.g., US and Canada), excluding countries that were separated by another nation (e.g., Canada and Mexico) and/or the high seas (e.g., Canada and Spain). Consequently, our results likely provide a conservative estimate of transboundary species, as we did not cover all marine taxa in the world [@Reygondeau:2019uh] and only analyzed species present in all four data sources. Nevertheless, our results are representative of a substantial proportion of the world's marine catches and revenue from economically important marine fisheries. Secondly, the predicted species' distribution is affected by the uncertainties of the environmental data and models used for such predictions. Structural differences within Earth System Models result in variations of oceanic conditions, which undoubtedly affect the ENM-Nereus. We averaged results from the three models in an effort to capture the structural variation across models. Natural climate variability is a major driver of marine species' distributions, potentially removing a species from an EEZ for long periods (e.g. Anchovies and Sardines "regimes" in the Eastern Pacific are strongly influenced by water temperature decal oscillations [@Chavez:2003jr]). Thus, a species’ distribution can potentially be reduced or shifted in such a way that it only covers one EEZ until oceanic conditions change again and the associated species' distribution expands. To account for such climate variability, we derived the ENM-Nereus results as an average of oceanic conditions between 1970 and 2000. This is not an issue for the SDM-SAU as it does not directly require environmental variables [@Pauly:2016uz]. Thirdly, we assumed that, if the *Sea Around Us* reconstructed data recorded a species as caught in any given grid cell, then the species was present within that grid cell. While catch data are not exempt of uncertainty, in most cases, differences between the*Sea Around Us* and the FAO self-reported data are smaller towards the end of the time series [@Pauly:2019bf]. Thus, we limited the catch data reference period in our analyses to between 2005 and 2014, the last ten years of data available. Likewise, the spatial allocation of the catch is subjected to imprecision, mainly produced by differences in the spatial scale of the original data and the method employed by the *Sea Around Us*. Finally, this study’s results only considered species for which all datasets agreed on presence and had and *Area Index* of at least 25% (e.g., the species shared distribution was at least 25% in each EEZ). Therefore, again, our approach presents a relatively conservative estimate of the number of transboundary species in the world. Using a smaller *Area Index* will result in more transboundary species (Supplementary Fig. S2).

\clearpage

# References

>>>>

---
output:
  bookdown::pdf_book:
    toc : no
    keep_tex: no
    number_sections: yes
    fig_width: 6 
    fig_height: 4 
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ../References/FishForVisa_Ref.bib
csl: ../References/nature_cc.csl
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
---

@@

## Introduction

@@

### Materials and Methods

### Databases of species' geographic distribution

The *Sea Around Us* has reconstructed global fisheries catches from 1950 to 2014, identifying commercial marine fish and invertebrates as well as fishing regions [@Zeller2016]. We used the *Sea Around Us* reconstructed species list to determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs. The total number of species analyzed was 938, representing 67% of identified species by the *Sea Around Us* and accounting for 96.52% of the catch identified at the species’ level. To determine the current distribution of exploitable marine species, we used four data sources of species-distributions: (*i*) occurrence data, (*ii*) an Ensemble Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (Table @\ref(tab:Data_Table)). Each source represents a different method of estimating the distribution of a given species, thus providing a more robust result than an analysis focused on a single data source. Only commercial fished species with data from all four sources were included in our analysis.

@@

data.frame(Source,
           Abreviation,
           Main_Method,
           Sources) %>% 
  rename("Main Method"=Main_Method) %>% 
  kable("latex", booktabs = T,
        caption = "Summary of data sources to estimate species' distributions. All data has a ressolution of 0.5 degrees latitude x 0.5 degrees longitude. All data is publically available, see references for details.") %>%
  kable_styling(full_width = T) %>%
  column_spec(2,width = "2cm") %>% #wraps text in column2
  column_spec(3,width = "5cm") #%>% #wraps text in column2
  # write_csv("Table_1.csv")
```

#### *Occurrence data*

@@

#### Criteria 3; Spatial distribution

Finally, to have a more robust result and not categorize a species as transboundary based on its presence in a single $0.5^{\circ}$ x $0.5^{\circ}$ grid cell within an EEZ, we computed an *Area Index*. The *Area Index* represents the proportion of a given species' overarching shared distribution between neighboring EEZs accounted for by each individual EEZ. We classified a species as transboundary if both neighboring EEZs enclosed over 25% of the species joint shared distribution. While a species that has less than the selected threshold is not considered transboundary in this paper, this threshold can be lowered for a more relaxed result or increased for a more conservative estimate (Figure \@(fig:Call-Sup-2)).

### Fisheries trends

@@

We used catch data as described above to determine the catch trend of each species within an EEZ. Although this method has previously been used to estimate stock status [@Grainger:1996tk], the categories presented here are intended to represent catch trends, and not fishing status for each species as many environmental and social-economic factors (e.g., temperature, markets, fishing policies, and fishing effort) affect catches [@Pauly:2013gs; @Branch:2008kk]. We only assessed species within each EEZ for which at least ten years of data were available between 1951 and 2014 and with at least five consecutive years of data. Three final categories were drawn for each species depending on catch volume within each EEZ (e.g., present, maximum, and minimum EEZ's historical catch) and the year (e.g., year of maximum historical catch of the species within that EEZ) (Table Table \@ref(tab:CallTab2)) [@Kleisner:2011wn]. Accordingly, Category A represents fisheries that are registering increases in catch (“increasing”); Category B, species that have a constant catch rate (“constant”); and Category C, species that have registered declines in catch over the last 10 years (“decreasing”). Finally, we reported the predominant category over the time period considered.

<!-- Table 2. Rules to determine the category of each analyzed species. -->

```{r CallTab2, eval = T, echo = F}

@@

data.frame(Categories,
           Rules) %>% 
  kable("latex", 
        booktabs = T,
        caption = "Rules to determine the category of each transboundary species") %>%
  kable_styling(full_width = T) %>%
  column_spec(1,width = "3cm") %>% #wraps text in column2
  column_spec(2,width = "13cm") #%>% #wraps text in column2
  # write_csv("Table_2.csv")

@@

We have identified four key uncertainties in the method utilized that may affect the estimation of transboundary species. Firstly, as we ran the analysis at the species level due to limited spatial-specific data on species sub populations (stocks), we are not able to identify transboundary stocks within EEZs. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level. Also, recent research suggests strong connectivity between stocks, even when separated by thousands of kilometers [@Ramesh:2019va] providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca]. **However, it is important to acknowledge that there is considerable uncertainty associated in determining levels of connectivity across time and space for marine populations at different life stages, from larvae [@Kaplan:2016hs] to adults [@Archambault:2016ie]. This is of special concern, but-not-limited-to [@Archambault:2016ie], highly migratory species whose ranges span multiple jurisdictions and the high seas, such as tunas [@Moore:2020cd], billfishes [@Sepulveda:2019hq] and sharks [@Vandeperre:2014jk] challenging management decisions based on meta-populations [@Moore:2020cd].** Here, we limited the definition of “transboundary” to include species spanning only adjacent countries (e.g., US and Canada), excluding countries that were separated by another nation (e.g., Canada and Mexico) and/or the high seas (e.g., Canada and Spain). Consequently, our results likely provide a conservative estimate of transboundary species, as we did not cover all marine taxa in the world [@Reygondeau:2019uh] and only analyzed species present in all four data sources. Nevertheless, our results are representative of a substantial proportion of the world's marine catches and revenue from economically important marine fisheries. Secondly, the predicted species' distribution is affected by the uncertainties of the environmental data and models used for such predictions. Structural differences within Earth System Models result in variations of oceanic conditions, which undoubtedly affect the ENM-Nereus. We averaged results from the three models in an effort to capture the structural variation across models. Natural climate variability is a major driver of marine species' distributions, potentially removing a species from an EEZ for long periods (e.g. Anchovies and Sardines "regimes" in the Eastern Pacific are strongly influenced by water temperature decal oscillations [@Chavez:2003jr]). Thus, a species’ distribution can potentially be reduced or shifted in such a way that it only covers one EEZ until oceanic conditions change again and the associated species' distribution expands. To account for such climate variability, we derived the ENM-Nereus results as an average of oceanic conditions between 1970 and 2000. This is not an issue for the SDM-SAU as it does not directly require environmental variables [@Pauly:2016uz]. Thirdly, we assumed that, if the *Sea Around Us* reconstructed data recorded a species as caught in any given grid cell, then the species was present within that grid cell. While catch data are not exempt of uncertainty, in most cases, differences between the*Sea Around Us* and the FAO self-reported data are smaller towards the end of the time series [@Pauly:2019bf]. Thus, we limited the catch data reference period in our analyses to between 2005 and 2014, the last ten years of data available. Likewise, the spatial allocation of the catch is subjected to imprecision, mainly produced by differences in the spatial scale of the original data and the method employed by the *Sea Around Us*. Finally, this study’s results only considered species for which all datasets agreed on presence and had and *Area Index* of at least 25% (e.g., the species shared distribution was at least 25% in each EEZ). Therefore, again, our approach presents a relatively conservative estimate of the number of transboundary species in the world. Using a smaller *Area Index* will result in more transboundary species (Figure \@ref(fig:CallSup2).

## Results and Discussion

We identified 633 exploited transboundary species worldwide (67.5% of the 938 species analyzed), almost double previous estimates [@Teh:2015gd]. Between 2005 and 2014, national fleets targeting these transboundary species within EEZs caught an annual average of 48.5 million tonnes, representing 82.3% of EEZ-derived catches reconstructed by the *Sea Around us* at the species level (Figure \@ref(fig:Figure1 a)). These catches generated a yearly average of USD 77,591 million in fishing revenue (78.5% of the global fishing revenue) over the same time period. Our findings are considerably higher than the previous estimates of 34.2 million tonnes in catches and USD 40,187 million (in 2019 value) in fishing revenue from shared stocks [@Teh:2015gd].When we re-estimated the transboundary species’ catches using data consistent with those used in previous studies (FAO global reported data in year 2006), the 633 transboundary species identified here accounted for 40.4 million tonnes of annual catches (i.e., 18% higher than previously estimated). Our results suggests that the contribution of transboundary species to global catch and fishing revenue might have been previously underestimated due to an incomplete understanding of the transboundary nature of marine fished species. The 305 non-transboundary species (termed as ‘discrete’ species here, see methods - Determining transboundary species trait) accounted for a much smaller proportion of total catch and revenue; 2.8 million tonnes and USD 4,282 million annually respectively, on average, between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

```{r Figure1, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species and their contribution to global fisheries catch and revenue. a) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. b) Contribution of transboundary species to regional revenue (left) and catch (right). Regions classified according to the United Nations sub-regions. Points = mean +- sd. Revenue in 2010 real USD"}

<!-- ![Figure 1. Number of transboundary species and their contribution to global fisheries catch and revenue. a) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. b) Contribution of transboundary species to regional revenue (left) and catch (right). Regions classified according to the United Nations sub-regions: E. = Eastern, N. = Northern, S. = Southern, W. = Western, Ltn. Ame. & the Car. = Latin America and the Caribbean. Points = regional  mean ± sd. Revenue in 2019 real USD. Figure created using R version 3.5.2. Land and EEZ shapefiles from Natural Earth (www.naturalearthdata.com) and Sea Around Us (www.seaaroundus.org), respectively.](../Figures/Fig1.png) -->

At a regional level, we found that transboundary species are particularly economically important for Northern America (average country revenue = USD 4,680 $\pm$ 6,000 million) and Eastern Asia (USD 3,779 $\pm$ 3,093 million) (Figure \@ref(fig:Figure1 b)). The estimated per country revenues from transboundary species in these two regions is significantly different from other regions (one-way ANOVA; *DF*(16,165) = 5.081, *p* < 0.001, $\alpha$ = 0.05). China (USD 7,284 million) and the USA (USD 11,604 million) contribute 55% and 82% to the annual average revenue from 2005 to 2014  in Eastern Asia and Northern America, respectively. In addition to China and the USA, Russia (USD 7,379 million), Peru (USD 6,044 million) and Japan (USD 3,907 million) were among the top five countries with the most fishing revenue generated from transboundary species between 2005 and 2014 (Figure \@ref(fig:Figure1 a)). These five nations were responsible for 41% of the yearly global fisheries revenue from transboundary species.

Peru and Russia, having the two largest fisheries by total catch in the world [@FAO:2018KK], each responsible for over 5.8 million tonnes annually, on average, between 2005 and 2014  (Supplementary Fig. S1). Peru's catches consist mainly of Peruvian anchoveta (*Engraulis ringens*) and account for 79% of the national transboundary species production. Peru and Chile recently signed an agreement to work towards standardized stock assessments through coordinated management of the southern anchoveta stock [@UNDP:2016wp]. A similar management agreement was signed by Russia, Japan and the USA over shared Alaskan pollock (*Theragra chalcogramma*) in the Bering Sea in 1988 [@NOAA:gX6CtUTk]. Transboundary species also make large contributions to fisheries in Eastern Asia (one-way ANOVA; *Df*(16,158) = 2.265, *p* = 0.005, $\alpha$ = 0.05). China, the world's top fish producer [@FAO:2018KK], obtains one third (5.1 million tonnes) of its total catches from transboundary species, followed by Japan (1.8 million tonnes) and South Korea (1.06 million tonnes). Differences in the regional importance of transboundary fisheries can also be illustrated in terms of catch-revenue over the area ($km^2$) of the EEZ (Figure \@ref(fig:Figure2)). As an example, the aggregated EEZ area for all Northern European countries where transboundary species are present was 3.3 million $km^2$, the $6^{th}$ smallest of the 17 groups analyzed (Supplementary Table S1). However, it had the second highest average revenue (USD 26.1 thousand per $km^2$) and the highest average catch (19.9 tonnes per $km^2$) of transboundary species per EEZ area between 2005 and 2014. At the country level, countries from Western Europe accrued significantly more revenue from transboundary fisheries per $km^2$ than any other country (one-way ANOVA, *Df*(16,165) = 3.267, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.05; Figure \@ref(fig:Figure2)).

```{r Figure_2, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap=""}

<!-- ![Figure 2. Weighted benefits of transboundary species by $km^{2}$ and UN sub-region. a) Revenue in thousand 2019 USD. b) Catch in tonnes. Points are sub-region averages $\pm$ s.d. by country. Figure created using R version 3.5.2. Abbreviations as in Figure 1.](../Figures/Fig2.png) -->

We determined the catch trend of each species within each EEZ, classifying them as increasing (Category A), constant (Category B) or decreasing (Category C) (Figure \@ref(fig:Figure3)). While previous studies have demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Our intention here is to examine where the catch trends of transboundary species differ from non-transboundary species between 2005 and 2010 relative to historic catch since 1951 (See methods). We found significant differences for all catch trend categories for transboundary species (one-way ANOVA, *Df*(2,459) = 47.94, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.001), and no significant differences in catch trend categories for discrete species (one-way ANOVA, *Df*(2,106) = 1.885, *p* = 0.157; $\alpha$ = 0.05). We also found significant differences in catch trends when directly comparing transboundary to discrete species categories (MANOVA, *Df*(2,459) = 19.001, *p* < 0.001). Overall, transboundary species only targeted by one country are generally less likely to have a decreasing catch trend compared to instances where the shared species is fished by neighboring countries (Supplementary Table S2).

```{r FIgure_3, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species by catch category. Category A, increasing catch; Category B, constant catch; Category C, decreasing catch"}
